---
title: "Memory usage of dataframe columns"
output: github_document
---


Illustrative data: `starwars`
---

The examples below make use of the `starwars` and `storms` data from the `dplyr` package

```{r}
# some example data
data(starwars, package = "dplyr")
data(storms, package = "dplyr")
```

For illustrating comparisons of dataframes, use the `starwars` data and produce two new dataframes `star_1` and `star_2` that randomly sample the rows of the original and drop a couple of columns.

---
title: "Missingness and counting NAs"
output: github_document
---


```{r, message=FALSE, warning=FALSE}
library(dplyr)
star_1 <- starwars %>% sample_n(50)
star_2 <- starwars %>% sample_n(50) %>% select(-1, -2)
```

Illustrative data: `starwars`
---

The examples below make use of the `starwars` and `storms` data from the `dplyr` package

```{r}
# some example data
data(starwars, package = "dplyr")
data(storms, package = "dplyr")
```

For illustrating comparisons of dataframes, use the `starwars` data and produce two new dataframes `star_1` and `star_2` that randomly sample the rows of the original and drop a couple of columns.

```{r, message=FALSE, warning=FALSE}
library(dplyr)
star_1 <- starwars %>% sample_n(50)
star_2 <- starwars %>% sample_n(50) %>% select(-1, -2)
```




`inspect_mem()` for a single dataframe
---

To explore the memory usage of the columns in a data frame, use `inspect_mem()`.  The command returns a `tibble` containing the size of each column in the dataframe.  

```{r}
library(inspectdf)
inspect_mem(starwars)
```

A barplot can be produced by passing the result to `show_plot()`:

```{r}
inspect_mem(starwars) %>% show_plot()
```

`inspect_mem()` for two dataframes
---

When a second dataframe is provided, `inspect_mem()` will create a dataframe comparing the size of each column for both input dataframes.  The summaries for the first and second dataframes are show in columns with names appended with `_1` and `_2`, respectively.

```{r}
inspect_mem(star_1, star_2)
```

```{r}
inspect_mem(star_1, star_2) %>% show_plot()
```

