---
title: "Correlation diagnostics for numeric columns"
output: github_document
---

Illustrative data: `starwars`
---

The examples below make use of the `starwars` and `storms` data from the `dplyr` package

```{r}
# some example data
data(starwars, package = "dplyr")
data(storms, package = "dplyr")
```

For illustrating comparisons of dataframes, use the `starwars` data and produce two new dataframes `star_1` and `star_2` that randomly sample the rows of the original and drop a couple of columns.

```{r, message=FALSE, warning=FALSE}
library(dplyr)
star_1 <- starwars %>% sample_n(50)
star_2 <- starwars %>% sample_n(50) %>% select(-1, -2)
```


`inspect_cor()` for a single dataframe
---

`inspect_cor()` returns a tibble containing Pearson's correlation coefficient, confidence intervals and $p$-values for pairs of numeric columns .  The function combines the functionality of `cor()` and `cor.test()` in a more convenient wrapper.  

```{r}
library(inspectdf)
inspect_cor(storms)
```

A plot showing point estimate and confidence intervals is printed when using the `show_plot()` function.  Note that intervals that straddle the null value of 0 are shown in gray:

```{r}
inspect_cor(storms) %>% show_plot()
```

Notes:

+ The tibble is sorted in descending order of the absolute coefficient $|\rho|$.
+ `inspect_cor` drops missing values prior to calculation of each correlation coefficient.  
+ The `p_value` is associated with the null hypothesis $H_0: \rho = 0$.


`inspect_cor()` for two dataframes
---

When a second dataframe is provided, `inspect_cor()` returns a tibble that compares correlation coefficients of the first dataframe to those in the second.  The `p_value` column contains a measure of evidence for whether the two correlation coefficients are equal or not.  

```{r}
inspect_cor(storms, storms[-c(1:200), ])
```

To plot the comparison of the top 20 correlation coefficients: 

```{r}
inspect_cor(storms, storms[-c(1:200), ]) %>% 
  slice(1:20) %>%
  show_plot()
```

Notes:

+ Smaller `p_value` indicates stronger evidence against the null hypothesis $H_0: \rho_1 = \rho_2$ and an indication that the true correlation coefficients differ.
+ The visualisation illustrates the significance of the difference using a coloured bar underlay.  Coloured bars indicate evidence of inequality of correlations, while gray bars indicate equality.  
+ For a pair of features, if either coefficient is `NA`, the comparison is omitted from the visualisation.
+ The significance level can be specified using the `alpha` argument to `inspect_cor()`.  The default is `alpha = 0.05`.

