---
title: "Missingness and counting NAs"
output: github_document
---

<!-- # ```{r, echo = FALSE} -->
<!-- # knitr::opts_chunk$set( -->
<!-- #   fig.path = "/vignettes/pkgdown/man/figures/inspect_na-" -->
<!-- # ) -->
<!-- # ``` -->
<!-- #  -->


Illustrative data: `starwars`
---

The examples below make use of the `starwars` and `storms` data from the `dplyr` package

```{r}
# some example data
data(starwars, package = "dplyr")
data(storms, package = "dplyr")
```

For illustrating comparisons of dataframes, use the `starwars` data and produce two new dataframes `star_1` and `star_2` that randomly sample the rows of the original and drop a couple of columns.

```{r, message=FALSE, warning=FALSE}
library(dplyr)
star_1 <- starwars %>% sample_n(50)
star_2 <- starwars %>% sample_n(50) %>% select(-1, -2)
```



`inspect_na()` for a single dataframe
---

`inspect_na()` summarises the prevalence of missing values by each column in a data frame.  A tibble containing the count (`cnt`) and the overall percentage (`pcnt`) of missing values is returned.

```{r}
library(inspectdf)
inspect_na(starwars)
```

A barplot can be produced by passing the result to `show_plot()`:

```{r}
inspect_na(starwars) %>% show_plot()
```

`inspect_na()` for two dataframes
---

When a second dataframe is provided, `inspect_na()` returns a tibble containing counts and percentage missingness by column, with summaries for the first and second data frames are show in columns with names appended with `_1` and `_2`, respectively.  In addition, a $p$-value is calculated which provides a measure of evidence of whether the difference in missing values is significantly different.

```{r}
inspect_na(star_1, star_2)
```

```{r}
inspect_na(star_1, star_2) %>% show_plot()
```

Notes:

+ Smaller $p$-values indicate stronger evidence of a difference in the missingness rate for a single column
+ If a column appears in one data frame and not the other - for example `height` appears in `star_1` but nor `star_2`, then the corresponding `pcnt_`, `cnt_` and `p_value` columns will contain `NA`
+ Where the missingness is identically 0, the `p_value` is `NA`.
+ The visualisation illustrates the significance of the difference using a coloured bar overlay.  Orange bars indicate evidence of equality or missingness, while blue bars indicate inequality.  If a `p_value` cannot be calculated, no coloured bar is shown.
+ The significance level can be specified using the `alpha` argument to `inspect_na()`.  The default is `alpha = 0.05`.
