---
title: "Exploring and visualising categorical features"
output: github_document
---

## Illustrative data: `starwars`

The examples below make use of the `starwars` and `storms` data from the `dplyr` package

```{r}
# some example data
data(starwars, package = "dplyr")
data(storms, package = "dplyr")
```

For illustrating comparisons of dataframes, use the `starwars` data and produce two new dataframes `star_1` and `star_2` that randomly sample the rows of the original and drop a couple of columns.

```{r, message=FALSE, warning=FALSE}
library(dplyr)
star_1 <- starwars %>% sample_n(50)
star_2 <- starwars %>% sample_n(50) %>% select(-1, -2)
```

## `inspect_cat()` for a single dataframe

`inspect_cat()` returns a tibble summarising categorical features in a data frame, combining the functionality of the `inspect_imb()` and `table()` functions.  The tibble generated contains the columns  

+ `col_name` name of each categorical column
+ `cnt` the number of unique levels in the feature
+ `common` the most common level (see also `inspect_imb()`)  
+ `common_pcnt` the percentage occurrence of the most dominant level  
+ `levels` a list of tibbles each containing frequency tabulations of all levels

```{r}
library(inspectdf)
inspect_cat(starwars)
```

For example, the levels for the `hair_color` column are

```{r}
inspect_cat(starwars)$levels$hair_color
```

Note that by default, if `NA` values are present, they are counted as a distinct categorical level.  A barplot is printed showing the relative split when passing the result to `show_plot()`:

```{r}
inspect_cat(starwars) %>% show_plot()
```

The argument `high_cardinality` in the `show_plot()` function can be used to bundle together categories that occur only a small number of times.  For example, to combine categories only occurring once, use:

```{r}
inspect_cat(starwars) %>% 
  show_plot(high_cardinality = 1)
```

The resulting bundles are shown in purple.

## `inspect_cat()` for two dataframes

When two dataframes are compared using `inspect_cat()`, list columns are returned for categorical columns common to both: `lvls_1` and `lvl2_2`.  In addition, the Jensen-Shannon divergence (`jsd`) and _p_ values associated with Fisher's exact test (`fisher_p`) are returned to enable comparison of the distribution of levels in each pair of columns.

```{r}
inspect_cat(star_1, star_2)
```

```{r}
inspect_cat(star_1, star_2) %>% show_plot()
```


