---
title: "Feature imbalance for categorical columns"
output: github_document
---


Illustrative data: `starwars`
---

The examples below make use of the `starwars` and `storms` data from the `dplyr` package

```{r}
# some example data
data(starwars, package = "dplyr")
data(storms, package = "dplyr")
```

For illustrating comparisons of dataframes, use the `starwars` data and produce two new dataframes `star_1` and `star_2` that randomly sample the rows of the original and drop a couple of columns.

```{r, message=FALSE, warning=FALSE}
library(dplyr)
star_1 <- starwars %>% sample_n(50)
star_2 <- starwars %>% sample_n(50) %>% select(-1, -2)
```


`inspect_imb()` for a single dataframe
---

Understanding categorical columns that are dominated by a single level can be useful.  `inspect_imb()` returns a tibble containing categorical column names (`col_name`); the most frequently occurring categorical level in each column (`value`) and `pctn` & `cnt` the percentage and count which the value occurs.  The tibble is sorted in descending order of `pcnt`. 

```{r}
library(inspectdf)
inspect_imb(starwars)
```

 A barplot is printed by passing the result to the `show_plot()` function:
 
```{r}
inspect_imb(starwars) %>% show_plot()
```


`inspect_imb()` for two dataframes
---

When a second dataframe is provided, `inspect_imb()` returns a tibble that compares the frequency of the most common categorical values of the first dataframe to those in the second.  The `p_value` column contains a measure of evidence for whether the true frequencies are equal or not. 

```{r}
inspect_imb(star_1, star_2)
```


```{r}
inspect_imb(star_1, star_2) %>% show_plot()
```

+ Smaller `p_value` indicates stronger evidence against the null hypothesis that the true frequency of the most common values is the same.
+ The visualisation illustrates the significance of the difference using a coloured bar overlay.  Orange bars indicate evidence of equality of the imbalance, while blue bars indicate inequality.  If a `p_value` cannot be calculated, no coloured bar is shown.
+ The significance level can be specified using the `alpha` argument to `inspect_imb()`.  The default is `alpha = 0.05`.

