---
title: "Exploring and visualising categorical features"
output: github_document
---



Illustrative data: `starwars`
---

The examples below make use of the `starwars` from the `dplyr` package. 

```{r, message=FALSE, warning=FALSE}
library(dplyr)
data(starwars, package = "dplyr")

# print the first few rows
head(starwars)
```



## `inspect_cat()` for a single data frame

`inspect_cat()` returns a tibble summarising categorical features in a data frame, combining the functionality of the `inspect_imb()` and `table()` functions.  The tibble generated contains the columns  

+ `col_name` name of each categorical column
+ `cnt` the number of unique levels in the feature
+ `common` the most common level (see also `inspect_imb()`)  
+ `common_pcnt` the percentage occurrence of the most dominant level  
+ `levels` a list of tibbles each containing frequency tabulations of all levels

```{r}
library(inspectdf)

# explore the categorical features 
x <- inspect_cat(starwars)
x
```

For example, the levels for the `hair_color` column are

```{r}
# show frequency tibble for `hair_color` column:
x$levels$hair_color
```

Note that by default, if missing (`NA`) values are present, they are counted as a distinct categorical level.  A barplot showing the composition of each categorical column can be created using the `show_plot()` function.  Note how missing values are shown as grey bars:

```{r}
x %>% show_plot()
```

The argument `high_cardinality` in the `show_plot()` function can be used to bundle together categories that occur only a small number of times.  For example, to combine categories only occurring once, use:

```{r}
x %>% 
  show_plot(high_cardinality = 1)
```

The resulting bundles are shown in purple.

## `inspect_cat()` for two data frames

To illustrate the comparison of two data frames, we first create two new data frames by randomly sampling the rows of `starwars` and also dropping some of the columns.  The results are assigned to the objects `star_1` and `star_2`:

```{r}
# sample 50 rows from `starwars`
star_1 <- starwars %>% sample_n(50)
# sample 50 rows from `starwars` and drop the first two columns
star_2 <- starwars %>% sample_n(50) %>% select(-1, -2)
```


To compare the character columns in a pair of data frames, use the  `inspect_cat()`: 

```{r}
inspect_cat(star_1, star_2)
```

The tibble returned has the following columns

+ `jsd`, the Jensen-Shannon divergence: a measure of how different the distribution of levels are between columns with the same name present in both data frames.  Values are between 0 and 1 - values closer to 1 indicate bigger differences in distribution.
+ `pval`, _p_ values associated with a modified $\chi^2$ test of the relative frequencies of levels in columns with the same name present in both data frames.  
+ `lvls_1` and `lvl2_2` are named list columns containing the frequency tables for each column in the first and second data frame input to `inspect_cat()`

<!-- # ```{r} -->
<!-- # inspect_cat(star_1, star_2) %>% show_plot() -->
<!-- # ``` -->


